Empowering (Domain) Recursion in Symmetric Weighted First-Order Model Counting

### 2021-10-17 talk/talk.tex
Make the greedy search code more flexible in preparation for breadth-first search
A barebones implementation of breadth-first search

\author{Paulius Dilkas}
\title{Symmetric Weighted First-Order Model Counting and Factorial-Like Functions}
\date{}

@@

\begin{frame}{Counting (Unweighted) Functions: Currently Unliftable (1)}
  \cite{stanley2011enumerative}
  \begin{block}{Functions $M \to N$ (let $|M| = m$ and $|N| = n$)}
    Theory:
    \begin{align*}
      \forall x \in M. \forall y, z \in N. &P(x, y) \land P(x, z) \Rightarrow y=z \\
      \forall x \in M. \exists y \in N. &P(x, y)
    \end{align*}
    Answer: $n^m$.
  \end{block}
  \begin{block}{Injections}
    Theory:
    \begin{align*}
      \forall x \in M. \forall y, z \in N. &P(x, y) \land P(x, z) \Rightarrow y=z \\
      \forall x \in M. \exists y \in N. &P(x, y) \\
      \alert{\forall w, x \in M. \forall y \in N.} &\alert{P(w, y) \land P(x, y) \Rightarrow w=x}
    \end{align*}
    Answer: $n^{\underline{m}} = n \cdot (n-1)\cdots(n-m+1)$ if $m \le n$ and $0$ otherwise (for positive $m$ and $n$).
  \end{block}
\end{frame}

@@

>>>>

\author{Paulius Dilkas}
\title{Empowering (Domain) Recursion in Symmetric Weighted First-Order Model Counting}
\date{}

@@

\begin{frame}{Counting (Unweighted) Functions: Currently Unliftable (1)}
  %% \cite{stanley2011enumerative}
  %% \begin{block}{Functions $M \to N$ (let $|M| = m$ and $|N| = n$)}
  %%   Theory:
  %%   \begin{align*}
  %%     \forall x \in M. \forall y, z \in N. &P(x, y) \land P(x, z) \Rightarrow y=z \\
  %%     \forall x \in M. \exists y \in N. &P(x, y)
  %%   \end{align*}
  %%   Answer: $n^m$.
  %% \end{block}
  \begin{block}{Injections}
    Theory:
    \begin{align*}
      \forall x \in M. \forall y, z \in N. &P(x, y) \land P(x, z) \Rightarrow y=z \\
      \forall x \in M. \exists y \in N. &P(x, y) \\
      \forall w, x \in M. \forall y \in N. &P(w, y) \land P(x, y) \Rightarrow w=x
    \end{align*}
    Answer: $n^{\underline{m}} = n \cdot (n-1)\cdots(n-m+1)$ if $m \le n$ and $0$ otherwise (for positive $m$ and $n$).

    \pause
    \alert{Answer found by Forclift:}
    \[
    f(m, n) = \sum_{k=0}^m \binom{m}{k}(-1)^{m-k}g(k, n),
    \]
    where
    \[
    g(k, n) =
    \begin{cases}
      1 & \text{if } k = 0 \\
      \sum_{l=0}^n [l < 2]g(k-1, n-l) & \text{otherwise.}
    \end{cases}
    \]
    (exponential...)
  \end{block}
\end{frame}

@@

\begin{frame}{Domain Size Comparisons}
  \begin{block}{$|N| < 2$ (Forclift can already do this!)}
    \[
    \phi \equiv \forall x, y \in N \text{ s.t. } x \ne y\text{, } P,
    \]
    where $w(P) = 0$. The algebraic (multiplicative) contribution of $\phi$ is $w(P)^{\mathrm{gr}(\phi)}$.
    \begin{itemize}
    \item If $|N| \ge 2$, then $\mathrm{gr}(\phi) > 0$, and so $0^{\mathrm{gr}(\phi)} = 0$.
    \item If $|N| < 2$, then we get $0^0 = 1$.
    \item Of course, 2 can be replaced by any other positive integer.
    \end{itemize}
  \end{block}
  \pause
  \begin{block}{$m < n$ (possible, but a bit complicated)}
    \[
    f(m, n) = \begin{cases}
      0 & \text{if } n = 0 \\
      1 & \text{if } m = 0 \text{ and } n > 0 \\
      f(m-1, n-1) & \text{otherwise.}
    \end{cases}
    \]
  \end{block}
\end{frame}
